Variational Monte Carlo on Bosonic Systems

### 2018-04-04 project_1/article/article.tex
Add first draft of MC integration

        To get around\footnote{Not exactly \emph{get around} per se, but make
        the problem solvable at all.} the trouble of evaluating the
        multidimensional-integral we employ the method of Monte Carlo
        integration.

@@

        When we are performing the Monte Carlo sampling we are interested in the
        expected value of the local energy as this will serve as our estimate of
        the true energy. Mathematically this is expressed as
        \begin{align}
            \expval{E_L}
            &=
            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T{\vf{r}}}
            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
            =
            \frac{
                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L
            }{
                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
            },
        \end{align}
        where we have multiplied and divided by the trial wavefunction to get
        the latter expression.

>>>>

        To get around\footnote{Not exactly \emph{get around} per se, but make
        the problem solvable at all.} the trouble of evaluating the
        multidimensional-integral we employ the method of Monte Carlo
        integration. This consists of approximating an integral of the
        expectation value of a certain quantitity by a sum.
        \begin{align}
            \expval{x}
            &=
            \int_{\mathbb{R}}xp(x)\dd x
            \approx
            \frac{1}{N}\sum_{i = 1}^N x_ip(x_i),
        \end{align}
        where $x$ is a random variable with a probability density function
        $p(x)$ and $\{x_i\}$ for $i \in \{1, \dots, N\}$ is the set of drawn
        random variables. Here $N$ is the number of Monte Carlo samples chosen
        to approximate the integral. By the law of large numbers we expect that
        \begin{align}
            \lim_{N \to \infty}\frac{1}{N}\sum_{i = 1}^N x_ip(x_i)
            = \expval{x}.
        \end{align}

        For us to be able to take Monte Carlo integration to the quantum realm
        we have to construct a probability density function from the
        wavefunction. Luckily the wavefunction squared is just such a quantity.
        Given a normalized wavefunction $\ket{\psi}$ we can compute the
        expectation value of a quantum observable $\mathcal{O}$ using Monte
        Carlo integration.
        \begin{align}
            \expval{\mathcal{O}}
            &= \bra{\psi}\mathcal{O}\ket{\psi}
            =
            \int\dd\vf{r}
            \psi^{*}(\vf{r})\mathcal{O}\psi(\vf{r})
            \\
            &=
            \int\dd\vf{r}
            |\psi(\vf{r})|^2
            \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})},
        \end{align}
        where we have multiplied and divided by the wavefunction. We now define
        the probability density function, $p(\vf{r})$, as
        \begin{align}
            p(\vf{r}) = |\psi(\vf{r})|^2,
        \end{align}
        and the local quantum observable
        \begin{align}
            \mathcal{O}_L(\vf{r})
            &= \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
        \end{align}
        We see that if the chosen wavefunction is an eigenstate of the quantum
        operator, the local quantum observable becomes a constant.  We can thus
        approximate the expecatation value of the quantum observable using Monte
        Carlo integration.
        \begin{align}
            \expval{O}
            &=
            \int\dd\vf{r}
            p(\vf{r})
            \mathcal{O}_L(\vf{r})
            \approx
            \frac{1}{N}\sum_{i = 1}^N
            p(\vf{r}_i)\mathcal{O}_L(\vf{r}_i).
        \end{align}
        In the case of a non-normalized wavefunction we need to divide by a
        normalization factor in the probability density function.

@@

        When we are performing the Monte Carlo sampling we are interested in the
        expected value of the local energy as this will serve as our estimate of
        the true energy. Mathematically this is expressed as
        \begin{align}
            \expval{E_L}
            &=
            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
            =
            \frac{
                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L(\vf{r})
            }{
                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
            },
        \end{align}
        where we have multiplied and divided by the trial wavefunction to get
        the latter expression.
